Harnessing the cloud to understand light: A whitepaper detailing the \MaxwellFDS core solver

### 2012-08-29 fds/fds.tex
Started on the plan!

\documentclass{article}
\usepackage{pstricks}
\usepackage{amsmath}
\newcommand{\E}[2]{\begin{align}#2\label{eq:#1}\end{align}}
\newcommand{\EE}[2]{\begin{subequations}\begin{align}#2\end{align}\label{eq:#1}\end{subequations}}
\newcommand{\curl}{\nabla\times}
\newcommand{\eq}[1]{\eqref{eq:#1}}
\newcommand{\fig}[1]{figure~\ref{fig:#1}}
\newcommand{\MaxwellFDS}{MaxwellFDS }
\newcommand{\BI}{\begin{itemize}\item}
\newcommand{\EI}{\end{itemize}}
\newcommand{\I}{\item}
\newcommand{\pf}[1]{\frac{\partial}{\partial #1}} % d/d_
\newcommand{\prim}{\text{prim}}
\newcommand{\dual}{\text{dual}}
\newcommand{\mx}{\text{max}}
\newcommand{\diag}{\text{diag}}
\title{Harnessing the cloud to understand light: 
        A whitepaper detailing the \MaxwellFDS core solver}
\begin{document}
\maketitle
\tableofcontents

@@

\subsection{The advantages of frequency-domain solutions}
The advantages of solving Maxwell's equations in the  \emph{frequency}-domain 
    as opposed to simulating them in the \emph{time}-domain include
    \BI clean excitation of input modes,
    \I  precise characterization of material dispersion,
    \I  calculation of figure of merits outside of the simulation,
    \I  explicit measurement of simulation error, and
    \I  explicit calculation of eigenmodes. \EI

@@

@@

\subsection{Reserving instances}
\subsection{Use of spot instances}
Talk about pricing and how one can be kicked off.

\subsection{Web console for manual monitoring/shutdown}

@@

>>>>

\documentclass{article}
\usepackage{pstricks}
\usepackage{amsmath}
\input ../defs.tex
\newcommand{\MaxwellFDS}{MaxwellFDS }
\newcommand{\pf}[1]{\frac{\partial}{\partial #1}} % d/d_
\newcommand{\prim}{\text{prim}}
\newcommand{\dual}{\text{dual}}
\newcommand{\mx}{\text{max}}
\newcommand{\diag}{\text{diag}}
\title{Harnessing the cloud to understand light: 
        A whitepaper detailing the \MaxwellFDS core solver}
\begin{document}
\maketitle
\tableofcontents

@@

\subsection{The advantages of frequency-domain solutions}
The advantages of solving Maxwell's equations in the  \emph{frequency}-domain 
    as opposed to simulating them in the \emph{time}-domain include
    \BI clean excitation of input modes,
    \I  precise characterization of material dispersion,
    \I  calculation of figure of merits outside of the simulation,
    \I  using a variable-resolution rectangular grid,
    \I  explicit measurement of simulation error, and
    \I  explicit calculation of eigenmodes. \EI

@@

\subsubsection{Grid: efficient, variable-resolution rectangular grid}
Frequency-domain simulations allow for efficient variable-resolution
    rectangular grids which are not possible using time-domain simulations.

In time-domain simulations, changing the grid size imposes a large 
    computational penalty in that the time step of the entire simulation
    must be reduced in order to satisfy the Courant condition at the
    point of highest resolution.
In contrast, a frequency-domain simulation does not even need
    to consider the idea of a time-step, and 
    so is free from this constraint.

@@

\subsection{Instance type used}
\MaxwellFDS builds a cluster
    using Amazon's cluster GPU instances (\texttt{cg1.4xlarge}),
    which each contain two M2050 Nvidia Tesla GPUs.
Therefore, an $N$-node \MaxwellFDS cluster will contain $2N$ GPUs.

\MaxwellFDS is built on Nvidia Tesla technology in order 
    to take care of the immense processing power available
    for parallelizable problems.

Additionally, all of a cluster's instances are launched
    in a single placement group, % Ref!
    which allows for faster communication between nodes.

\subsection{Use of spot instances}
Because scientific workloads are almost always of a batch nature
    (e.g. many simulations run at once),
    \MaxwellFDS takes advantage of Amazone EC2 spot instances
    in order to drastically lower the cost of using \MaxwellFDS.
Spot instances cost much less than reserved instances,
    and can typically be requested in much larger numbers.

To request a spot instance, one must submit a bid price
    for the desired instance.
Amazon will then allocate the desired instance for the user
    as long as the bid price is above the market price.
If the market price exceeds the bid price,
    the instance will be shut down and the user will no longer be billed.

Although large fluctuations in market price are not common,
    \MaxwellFDS uses a large bid price (currently 2 USD) close to the
    reserved instance price for \texttt{cg1.4xlarge} 
    in order to ensure that nearly all spot instance bids will be successful.

\subsection{Web console for manual monitoring/shutdown}
Of course, there is no need to trust what \MaxwellFDS is doing behind the scenes
    because Amazon provides a web console for monitoring and 
    manually controlling all instances reserved and in use by the user.

The web console for ec2 can be accessed from \texttt{console.aws.amazon.com}.
Of particular interest will be ``spot instance requests'' and ``instances''.

@@

\subsubsection{Mathematical definition of simulation error}
From \eq{maxwell wave E discretized} we derive the mathematical expression for 
    the simulation error which is simply    
    \E  {maxwell error}
        {\text{error} = \|Ax-b\|_2 / \|b\|_2,}
    where $A = A_1 \diag(\mu^{-1}) A_2 - \omega^2 \diag(\epsilon)$.
